---
title: "Add On slides"
author: "Jan-Philipp Kolb"
date: "21 2 2020"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


---

## Data Frames

<!--
A data frame is more general than a matrix - different columns can have different modes (numeric, character, factor, etc.). This is similar to SAS and SPSS datasets.
-->

A data frame is a collection of vectors - different columns can have different modes (numeric, character, factor, etc.).

### Three example vectors

```{r}
d <- c(1,2,3,4)
e <- c("red", "white", "red", NA)
f <- c(TRUE,TRUE,TRUE,FALSE)
```


### Bind the example vectors together:

```{r}
mydata <- data.frame(d,e,f)
```

### Give the columns some names

```{r}
names(mydata) <- c("ID","Color","Passed") # variable names
```




---

## Identify the elements of a data frame

There are a variety of ways to identify the elements of a data frame .

```{r,eval=F}
myframe[3:5] # columns 3,4,5 of data frame
myframe[c("ID","Age")] # columns ID and Age from data frame
myframe$X1 # variable x1 in the data frame 
```





## [Matrices](https://www.statmethods.net/input/datatypes.html)

All columns in a matrix must have the same mode (numeric, character, etc.) and the same length. The general format is:

```{r,eval=F}
# generates 5 x 4 numeric matrix
y <- matrix(1:20, nrow=5,ncol=4)
```


- `byrow=TRUE` indicates that the matrix should be filled by rows. 
- `byrow=FALSE` - matrix should be filled by columns (the default). 

```{r}
# an example
cells <- c(1,26,24,68)
mymatrix <- matrix(cells, nrow=2, ncol=2, byrow=TRUE) 
```



---

## Matrix - dimnames


- `dimnames` provides optional labels for the columns and rows. 

```{r}
# another example
rnames <- c("R1", "R2")
cnames <- c("C1", "C2")
mymatrix <- matrix(cells, nrow=2, ncol=2, byrow=TRUE,
  dimnames=list(rnames, cnames)) 
```




---

## Matrices and subscripts

Identify rows, columns or elements using subscripts. 

```{r,eval=F}
x[,4] # 4th column of matrix
x[3,] # 3rd row of matrix
x[2:4,1:3] # rows 2,3,4 of columns 1,2,3 
```



---

## Lists

An ordered collection of objects (components). A list allows you to gather a variety of (possibly unrelated) objects under one name.

```{r,eval=F}
# example of a list with 4 components -
# a string, a numeric vector, a matrix, and a scaler
w <- list(name="Fred", mynumbers=a, mymatrix=y, age=5.3)

# example of a list containing two lists
v <- c(list1,list2)
```


Identify elements of a list using the [[ ]] convention.

```{r,eval=F}
mylist[[2]] # 2nd component of the list
mylist[["mynumbers"]] # component named mynumbers in list
```



