---
title: "Introduction to R"
subtitle: "Graphing Data Using GGPlot"
author: "Jan-Philipp Kolb"
date: "`r format(Sys.time(), '%d %B, %Y')`"
location: "GESIS, Mannheim, Germany"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "../workshop.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
 
```{r, include=FALSE}
knitr::opts_chunk$set(echo = T, message = FALSE,warning=F)
```
 


## GGplot vs Base

`ggplot` attempts to create a consistent framework for build graphs 'layer by layer' in R. 

### You construct a graph by specifying:

1. The data.
2. An aesthetic (e.g., colors, line styles, the coordinate system, etc). 
3. A graph "geometry" (e.g., boxplot, scatterplot, etc). This is where you specify the kind of graph you want.
4. Labels. The plot title, axis labels, etc. 

---

## A scatterplot 

Creating the graph object and specifying the dataset.

```{r}
library(ggplot2)
car.graph <- ggplot(mtcars)
```


- Specifying what aesthetics to use. 
- In this case, the coordinate system to use -- meaning the x-y axis. 
- This is sometimes also referred to as the "mapping" being used.


```{r}
car.graph <- car.graph + aes(wt, mpg)
```

---

## Specifying the plot

```{r}
## Specifying the plot "geometry" in this case, a scatter plot.
car.graph + geom_point()
```

---

## Without intermediate states

We can also create the plot without storing all of the intermediate states

```{r}
ggplot(mtcars) + aes(wt, mpg) + geom_point()
```

---

## Exercise: ggplot with the Ames Housing dataset

- Load the Ames Housing dataset

```{r}
ames_df <- AmesHousing::make_ames()
```

- Make a scatterplot with the variables sale price and lot area
- If you have that try to color the points according to MS Zoning

---

## What the result should look like

```{r, echo=T,fig.width=12, fig.height=4}
ggplot(ames_df) +
    aes(x=ames_df$Sale_Price, y=ames_df$Lot_Area) +
    geom_point(aes(col=as.factor(ames_df$MS_Zoning))) + 
    labs(title="Scatterplot", color="MS Zoning") +
    theme(legend.title=element_text(color="black")) 
```

---

## The themes in ggplot 



```{r, echo=T}
## Minimal theme
ggplot(mtcars) + aes(wt, mpg) + geom_point() + 
    theme_minimal()
```

---

## Themes


```{r,eval=F,echo=T}
## the default
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_gray() 
## Dark, usually not recommended.
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_dark() 
## Very traditional
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_classic()
## Use if you don't want an axis.
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_void() 
```



- There are a number of other built in themes, but you get the idea. 
- You can also create your own themes if one of the built in ones doesn't do
what you want.

---

## Colors

```{r, echo=T}
car.graph <- ggplot(mtcars) + aes(wt, mpg) + 
    geom_point(color="red")
car.graph
```

---

## point/line styles.

```{r, echo=T}
car.graph <- ggplot(mtcars) + aes(wt, mpg) + 
    geom_point(shape=21) + geom_line(linetype=2)
car.graph
```

---
 
## A boxplot
```{r, echo=T}

ggplot(mtcars) + aes(as.factor(gear), mpg) +geom_boxplot()

```

---

## Adding a title/changing labels

- geom_text() adds text to a plot. 

- geom_label() adds stuff to make the text easier to read (e.g., a box around the text).

-  labs() modifies your labels/title.

-  theme() lets us manipulate stuff like the inclusion of a legend, its position, etc.

---

## Adding a title/changing labels

```{r, echo=T}

ggplot(mtcars) + aes(as.factor(gear), mpg) +
    geom_boxplot() +
    geom_text(aes(label=as.factor(gear), col="red")) +
    geom_label(aes(label=as.factor(gear), col="red"))+
    labs(x="A different label than earlier.", title="A boxplot, with observations labeled according to their number of gears") +
    theme(legend.position="none")

```

---

## Saving a graph

```{r, echo=T, eval=F}

## Saving to pdf
ggsave(car.graph, file="car_graph.pdf")

## Saving to pdf, while specifying dimensions of plot
ggsave(car.graph, file="car_graph.pdf", width = 20, 
       height = 20, units = "cm")

## Saving to png
ggsave(car.graph, file="car_graph.png")

## Other formats are possible. 
# See ?ggsave for more information.

```

---

## A barplot

```{r, echo=T}
## Data + aesthetics + geometry.
ggplot(mtcars)+aes(gear)+geom_bar()
```

---

## A histogram

```{r, echo=T}

## Data + aesthetics + geometry.
ggplot(mtcars)+aes(mpg)+geom_histogram()

```

---

## A density plot

```{r, echo=T}
ggplot(mtcars)+aes(mpg)+geom_density()
```

---

## Exercise: Histogram and density plot

Using your data from the scatterplot exercise, produce a histogram for
cont.var.x, and a density plot for cont.var.y.

```{r, echo=FALSE, eval=FALSE}
ggplot(simulated.dataset) + geom_histogram()+
    aes(x=cont.var.x)

ggplot(simulated.dataset) + geom_density()+
    aes(x=cont.var.y)

```

---

## Links to read on

- [**Top 50 ggplot2 Visualizations**](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
- [R cookbook for graphs](http://www.cookbook-r.com/Graphs/)
- [ggplot2 as part of the tidyverse](https://ggplot2.tidyverse.org/)
- [ggplot2-graphics-workshop](https://github.com/aosmith16/ggplot2-graphics-workshop/)
- [ggplot2 command references](https://ggplot2.tidyverse.org/reference/)
