---
title: "Graphing Data Using GGPlot"
author: "Kolb / Kariuki"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 10pt
output:
  beamer_presentation: 
    fig_caption: no
    fig_height: 3
    fig_width: 5
    fonttheme: structuresmallcapsserif
    highlight: zenburn
    colortheme: whale
    theme: Warsaw
  pdf_document:
    toc: yes
  html_document:
    keep_md: yes
  slidy_presentation:
    keep_md: yes
---
 
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,warning=F)
```
 
 


## GGplot vs Base

ggplot attempts to create a consistent framework for build graphs "layer by layer" in R. 

You construct a graph by specifying:

1. The data.
2. An aesthetic (e.g., colors, line styles, the coordinate system, etc). 
3. A graph "geometry" (e.g., boxplot, scatterplot, etc). This is where you specify the kind of graph you want.
4. Labels. The plot title, axis labels, etc. 


## A scatterplot 

creating the graph object and specifying the dataset.

```{r, echo=T}
library(ggplot2)
car.graph <- ggplot(mtcars)
```


- Specifying what aesthetics to use. In this case, the coordinate
- system to use -- meaning the x-y axis. This is sometimes also
- referred to as the "mapping" being used.


```{r}
car.graph <- car.graph + aes(wt, mpg)
```


## Specifying the plot

```{r}
## Specifying the plot "geometry" in this case, a scatter plot.
car.graph <- car.graph + geom_point()

## Calling the graph object reproduces the plot.
car.graph

## We can also create the plot without storing all of the intermediate
## states
car.graph <- ggplot(mtcars) + aes(wt, mpg) + geom_point()

car.graph

```


```{r, echo=T, eval=F}

## We can save the plot using the ggsave function 
# (more detail on this later).
ggsave(car.graph, filename="myplot.pdf")

```


## Exercise: A scatterplot

Use the code below to generate a random dataset. Then plot the results
as a scatterplot, using ggplot2. Don't worry if your labels are
different than mine. I'll cover some of those details after the
exercise.

```{r, echo=T, eval=T}
set.seed(42)
cat.var <- c(rep.int(0, 100), rep.int(1, 100))
cont.var.x <- rnorm(n=100, mean=2, sd=4)
cont.var.y <- cont.var.x + cont.var.x*cat.var
simulated.dataset <- data.frame(cont.var.x, cont.var.y, cat.var)
```

## What the result should look like

```{r, echo=F}
ggplot(simulated.dataset) +
    aes(x=cont.var.x, cont.var.y) +
    geom_point(aes(col=as.factor(cat.var))) + 
    labs(title="Simulated scatterplot.", color="Our made up categories") +
    theme(legend.title=element_text(color="black")) + 
    theme(legend.text = element_text(size = 8, colour = "red"))
     

```



## Specifying different aesthetics

### Themes 
```{r, echo=T}
## Minimal theme
ggplot(mtcars) + aes(wt, mpg) + geom_point() + 
    theme_minimal()
```

## Themes

```{r,eval=F}
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_gray() ## the default
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_dark() ## Dark, usually not recommended.
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_classic() ## Very traditional
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_void() ## Use if you don't want an axis.
```

```{r,eval=F,echo=F}
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_gray() ## the default
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_dark() ## Dark, usually not recommended.
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_classic() ## Very traditional
ggplot(mtcars) + aes(wt, mpg) + geom_point() + theme_void() ## Use if you don't want an axis.
```



There are a number of other built in themes, but you get the idea. You
can also create your own themes if one of the built in ones doesn't do
what you want.

### colors

```{r, echo=T}
car.graph <- ggplot(mtcars) + aes(wt, mpg) + 
    geom_point(color="red")
car.graph
```



## point/line styles.

```{r, echo=T}
car.graph <- ggplot(mtcars) + aes(wt, mpg) + 
    geom_point(shape=21) + geom_line(linetype=2)
car.graph
```

 
## A boxplot
```{r, echo=T}

ggplot(mtcars) + aes(as.factor(gear), mpg) +geom_boxplot()

```

## Adding a title/changing labels

- geom_text() adds text to a plot. 

- geom_label() adds stuff to make the text easier to read (e.g., a box around the text).

-  labs() modifies your labels/title.

-  theme() lets us manipulate stuff like the inclusion of a legend, its position, etc.

## Adding a title/changing labels

```{r, echo=T}

ggplot(mtcars) + aes(as.factor(gear), mpg) +
    geom_boxplot() +
    geom_text(aes(label=as.factor(gear), col="red")) +
    geom_label(aes(label=as.factor(gear), col="red"))+
    labs(x="A different label than earlier.", title="A boxplot, with observations labeled according to their number of gears") +
    theme(legend.position="none")

```

## Saving a graph

```{r, echo=T, eval=F}

## Saving to pdf
ggsave(car.graph, file="car_graph.pdf")

## Saving to pdf, while specifying dimensions of plot
ggsave(car.graph, file="car_graph.pdf", width = 20, 
       height = 20, units = "cm")

## Saving to png
ggsave(car.graph, file="car_graph.png")

## Other formats are possible. 
# See ?ggsave for more information.

```

## A barplot
```{r, echo=T}
## Data + aesthetics + geometry.
ggplot(mtcars)+aes(gear)+geom_bar()
```

## A histogram

```{r, echo=T}

## Data + aesthetics + geometry.
ggplot(mtcars)+aes(mpg)+geom_histogram()

```

## A density plot

```{r, echo=T}
ggplot(mtcars)+aes(mpg)+geom_density()
```

## Exercise: Histogram and density plot

Using your data from the scatterplot exercise, produce a histogram for
cont.var.x, and a density plot for cont.var.y.

```{r, echo=FALSE, eval=FALSE}
ggplot(simulated.dataset) + geom_histogram()+
    aes(x=cont.var.x)

ggplot(simulated.dataset) + geom_density()+
    aes(x=cont.var.y)

```

## A network plot

This generally needs an additional library: ggraph.


```{r, echo=T, eval=F}
install.packages("ggraph")
```

```{r, echo=T}
library(igraph)
library(ggraph)
## Setting a "seed" so that we get the same random graph 
## each time the code is run.
set.seed(42)
## Generating a random graph
g <- barabasi.game(100, m=5)
## Plotting the graph, specifying edge color, 
# specifying node shape.
ggraph(g)+ geom_edge_link(aes(color="red"))+geom_node_point()
```


##

```{r}
## Storing the degree of each node so that 
# we can use it later.
V(g)$degree <- degree(g)
## Same plot as above, but with node size 
# based on degree.
ggraph(g)+ geom_edge_link(aes(color="red"))+geom_node_point(aes(size=degree))

```

